### 2020-01-09 nlp_models.tex
minor reorg
discussion of GPT2 added

\paragraph{GPT vs GPT2}

@@

\paragraph{Power Law Exponents}

@@

\paragraph{Correlation Flow}

>>>>

In the past two years, nearly 100 open source, pre-trained DNNs for Natural Language Processing (NLP) have emerged,
based on the revolutionary Transformer architecture, including variants of BERT, Transformer-XML, GPT, etc.
The Transformer architectures consist of blocks of Attention layers, containing of 2 large, Feed Forward (Linear)
weight matrics \cite{Attn2017}. In contrast to the smaller pre-Activation maps arising in Cond2D layers,
the Attention matrices ar

\paragraph{GPT vs GPT2}
Here, we use the \emph{WeightWatcher} tool to analyze the OpenAI GPT and GPT2 models, which gives us
the opportunity to analyze the effect of both training the same model with different size data sets,
These GPT models have generated significant media attention because of its remarkable ability to
 generate fake text, and it's potential misuse. 
The original GPT and later GPT model have the same architecture and number of layers, but
the original GPT was released having been trained on a deficient data set.  Recently,
 however, a much improved GPT2 model has been released, which is remarkably good at creating 
auto-genetated text.

\charles{more here ?}
We analyze GPT models deployed with the popular HuggingFace PyTorch library.
GPT has 12 layers, with 4 Multi-head Attention Blocksm, giving $48$ Layer Weight Matrices $\mathbf{W}$.
Each Block has 2 components, the Self Attention (attn) and the Projection (proj) matrics.  
The self-attention  matrices are larger, of dimension ($2304\times 768$) or ($3072\times 768$).
The projection layer concatenates the self-attention results into a vector (of dimension $768$).
This gives many large, typically sparse weight matrices, and they can be
 \emph{poorly correlated} when trained with insufficient data--as we shall see below.

Because GPT and GPT are trained on different data sets, the initial Embedding matrices differ in shape.
GPT  has an initial Token and Positional Embedding layers, of dimension
$(40478\times 768)$ and $(512\times 768)$, resp, whereas GPT2 has input Embeddings of shape
$(50257\times 768)$ and $(1024\times 768)$, resp.  Interstingly, they also have very spectral properties,
also shown below.

\charles{Since we show this table, maybe we should discuss GPT2 scaling models}

@@

\paragraph{Heavy Tailed Power Law Exponents in GPT and GPT2}

@@

\paragraph{Correlation Flow in GPT and GPT2}
